The Galactic Delay-Time Distribution of Type Ia Supernovae:\\ A Chemical Evolution Perspective

### 2023-07-25 src/tex/ms.tex
[showyourwork] overleaf sync

@@

\begin{figure}
    \centering
    \includegraphics{figures/dtd_greggio05.pdf}
    \caption{Analytical DTDs from \citet[][solid curves]{Greggio2005-AnalyticalRates} and close approximation simple functions (dashed curves). Some functions are presented with a constant multiplicative factor for visual clarity.}
    \label{fig:dtd-greggio05}
    \script{dtd_greggio05.py}
\end{figure}

@@

\subsection{Stellar Migration}
\label{sec:migration}

We fit a Gaussian to the distribution of $\Delta R = R_{\rm final} - R_{\rm initial}$ from the \texttt{h277} output in bins of $R_{\rm initial}$ and age. Each Gaussian is centered at 0 and we find that the scale $\sigma_{\Delta R}$ is best described by the function

\begin{equation}
    \sigma_{\Delta R} = 1.35\,{\rm kpc} \Big(\frac{R_{\rm form}}{8\,{\rm kpc}}\Big)^{0.61} \Big(\frac{\tau}{1\,{\rm Gyr}}\Big)^{0.33}
    \label{eq:radial-migration}
\end{equation}

\noindent where $\tau$ is the age of the star particle.

@@

@@

\subsection{The Star Formation History}

Figure \ref{fig:onezone-delay-taustar} shows results from one-zone models with a power-law DTD which differ in their assumed minimum SN Ia delay time ($t_D$) or star formation efficiency timescale ($\tau_*$). In [Fe/H]-[O/Fe] space, doubling or halving the minimum delay time produces similar results to doubling or halving the SFE timescale. A shorter minimum delay time means SNe Ia will start to explode sooner after the star formation burst, giving less time for CCSNe to explode at the plateau of the abundance track. Similarly, a longer SFE timescale means fewer massive stars will be produced by the initial star formation burst, resulting in fewer CCSNe before the first SNe Ia explode.

The minimum delay and SFE timescale have similar effects on the abundance tracks but very different effects on the MDFs. Altering the SFE timescale changes the stellar distribution of [Fe/H], especially at the low-metallicity end, as shown in the top panel of Figure \ref{fig:onezone-delay-taustar}. However, as a change in the SFE timescale affects both the number of high-mass and low-mass stars produced, it does not affect the distribution of [O/Fe], as shown in the right-hand panel of Figure \ref{fig:onezone-delay-taustar}. Meanwhile, a change in the minimum SN Ia delay time does not impact the total number of SNe which explode, so the distribution of [Fe/H] is largely unaffected, while it does result in a change in the number of high-$\alpha$ stars formed and therefore the distribution of [O/Fe].

\begin{figure}
    \centering
    \includegraphics{figures/onezone_delay_taustar.pdf}
    \caption{Similar to Figure \ref{fig:onezone-powerlaw-slope} but with varying minimum SN Ia delay time ($t_D$) and star formation efficiency timescale ($\tau_*$). Tracks with $\tau_*=2$ Gyr are represented in black, while a track with a shorter $\tau_*=1$ Gyr is shown in red and a track with a longer $\tau_*=4$ Gyr is in yellow. The dotted, solid, and dashed black curves represent a minimum delay time of 40 Myr, 80 Myr, and 160 Myr, respectively. A power-law DTD with $\alpha=-1.1$ is assumed in all models.}
    \label{fig:onezone-delay-taustar}
    \script{onezone_delay_taustar.py}
\end{figure}

@@

\begin{figure*}
    \centering
    \includegraphics{figures/onezone_greggio05_single.pdf}
    \includegraphics{figures/onezone_greggio05_double.pdf}
    \caption{Abundance tracks and distributions from one-zone models with the analytical DTDs from \citet{Greggio2005-AnalyticalRates}. \textit{Left:} a comparison between the analytical SD DTD (solid magenta curve), an exponential DTD with a 1.5 Gyr timescale (dashed pink curve), and a power-law DTD with a slope of -1.1 (dotted black curve). \textit{Right:} a comparison between WIDE and CLOSE prescriptions for an analytical DD DTD (solid blue and green curves, respectively), power-law DTDs with an initial plateau of 1 Gyr and 300 Myr (dashed cyan and yellow curves, respectively), and a power-law DTD with no plateau and a slope of -1.1 (dotted black curve).}
    \label{fig:onezone-greggio05}
    \script{onezone_greggio05.py}
\end{figure*}

@@

>>>>

Studies of galactic chemical evolution typically assume a single form for the SN Ia DTD. This is the first comprehensive look at the DTD in a model for GCE which includes the effects of radial migration.

Why radial migration is important for SNe Ia.

Summary of observational studies of the DTD. These studies all assume a particular form for the DTD and fit for the specific parameters. They do not quantitatively favor one form over another (e.g., power-law over exponential).

@@

@@

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.32\linewidth]{figures/onezone_powerlaw_slope.pdf}
    \includegraphics[width=0.32\linewidth]{figures/onezone_exponential_timescale.pdf}
    \includegraphics[width=0.32\linewidth]{figures/onezone_plateau_width.pdf}
    \\
    \includegraphics[width=0.49\linewidth]{figures/onezone_minimum_delay.pdf}
    \includegraphics[width=0.49\linewidth]{figures/onezone_dtd.pdf}
    \caption{}
    \label{fig:onezone-multipanel}
    \script{onezone_multipanel.py}
\end{figure*}

@@

% \subsection{The Star Formation History}

% Figure \ref{fig:onezone-delay-taustar} shows results from one-zone models with a power-law DTD which differ in their assumed minimum SN Ia delay time ($t_D$) or star formation efficiency timescale ($\tau_*$). In [Fe/H]-[O/Fe] space, doubling or halving the minimum delay time produces similar results to doubling or halving the SFE timescale. A shorter minimum delay time means SNe Ia will start to explode sooner after the star formation burst, giving less time for CCSNe to explode at the plateau of the abundance track. Similarly, a longer SFE timescale means fewer massive stars will be produced by the initial star formation burst, resulting in fewer CCSNe before the first SNe Ia explode.

% The minimum delay and SFE timescale have similar effects on the abundance tracks but very different effects on the MDFs. Altering the SFE timescale changes the stellar distribution of [Fe/H], especially at the low-metallicity end, as shown in the top panel of Figure \ref{fig:onezone-delay-taustar}. However, as a change in the SFE timescale affects both the number of high-mass and low-mass stars produced, it does not affect the distribution of [O/Fe], as shown in the right-hand panel of Figure \ref{fig:onezone-delay-taustar}. Meanwhile, a change in the minimum SN Ia delay time does not impact the total number of SNe which explode, so the distribution of [Fe/H] is largely unaffected, while it does result in a change in the number of high-$\alpha$ stars formed and therefore the distribution of [O/Fe].

% \begin{figure}
%     \centering
%     \includegraphics{figures/onezone_delay_taustar.pdf}
%     \caption{Similar to Figure \ref{fig:onezone-powerlaw-slope} but with varying minimum SN Ia delay time ($t_D$) and star formation efficiency timescale ($\tau_*$). Tracks with $\tau_*=2$ Gyr are represented in black, while a track with a shorter $\tau_*=1$ Gyr is shown in red and a track with a longer $\tau_*=4$ Gyr is in yellow. The dotted, solid, and dashed black curves represent a minimum delay time of 40 Myr, 80 Myr, and 160 Myr, respectively. A power-law DTD with $\alpha=-1.1$ is assumed in all models.}
%     \label{fig:onezone-delay-taustar}
%     \script{onezone_delay_taustar.py}
% \end{figure}

@@

@@

\begin{figure*}
    \centering
    \includegraphics{figures/dtd_analytical.pdf}
    \includegraphics{figures/onezone_analytical_dtd.pdf}
    \caption{\textit{Left:} Analytical DTDs from \citet[][solid curves]{Greggio2005-AnalyticalRates} and close approximation simple functions (dashed curves). Some functions are presented with a constant multiplicative factor for visual clarity. \textit{Right:} Abundance tracks and distributions from one-zone models with the analytical DTDs from \citet{Greggio2005-AnalyticalRates}. \textit{Left:} a comparison between the analytical SD DTD (solid magenta curve), an exponential DTD with a 1.5 Gyr timescale (dashed pink curve), and a power-law DTD with a slope of -1.1 (dotted black curve). \textit{Right:} a comparison between WIDE and CLOSE prescriptions for an analytical DD DTD (solid blue and green curves, respectively), power-law DTDs with an initial plateau of 1 Gyr and 300 Myr (dashed cyan and yellow curves, respectively), and a power-law DTD with no plateau and a slope of -1.1 (dotted black curve).}
    \label{fig:analytical-dtd}
    \script{analytical_dtd_twopanel.py}
\end{figure*}

\section{Stellar Migration}
\label{app:migration}

We fit a Gaussian to the distribution of $\Delta R = R_{\rm final} - R_{\rm initial}$ from the \texttt{h277} output in bins of $R_{\rm initial}$ and age. Each Gaussian is centered at 0 and we find that the scale $\sigma_{\Delta R}$ is best described by the function
\begin{equation}
    \sigma_{\Delta R} = 1.35\,{\rm kpc} \Big(\frac{R_{\rm form}}{8\,{\rm kpc}}\Big)^{0.61} \Big(\frac{\tau}{1\,{\rm Gyr}}\Big)^{0.33}
    \label{eq:radial-migration}
\end{equation}
where $\tau$ is the age of the star particle.
